---
title: "Qianru_Liao_lab2"
author: "Qianru Liao"
date: "10/25/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---


# Lab 2
```{r, echo = FALSE}
library(GISTools)
library(RColorBrewer)
library(tbart)
library(leaflet)
library(sp)
library(ggplot2)
# Load the data and change the working directory where the downloaded data ﬁle is located
load('EN.RData')
#getwd()
#do not need to change the working environment 
set.seed(12345)
```

# Question1
Make maps of the study area shaded by the population density and the daily variability of PM2.5 in the year of 2011. The two variables are respectively named as pop.den, pm25.std in the spatial polygon data frame (en.county). 

## Make map of population density
```{r}
en.county1<-spTransform(en.county, "+proj=longlat +datum=WGS84 +no_defs")

m <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(0,1,2,3,4,5,Inf)
pal<-colorBin("Set1", domain = en.county$pop.den, bins = bins)

pop.den_map <- m%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(en.county$pop.den),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
    addLegend(pal = pal, values = ~en.county$pop.den, opacity = 0.5, title = 'Population Density in Buffalo',
            position = "bottomright")
pop.den_map
```

Discussion of the map:

In most rural areas, the population density in red are lower than 1 and the place are located in the rural areas, which are far away from the Buffalo center. And the high population density in light yellow which are higher than 5, are concentrated in the central Buffalo. The population density gets higher with the decline of the distance between their areas and Buffalo center.

## Make map of the daily variability of PM 2.5 in 2011
```{r}
m <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(4,4.2,4.4,4.6,4.8,5,5.2,Inf)
pal<-colorBin("Spectral", domain = en.county$pm25.std, bins = bins)

pm.25_map <- m%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(en.county$pm25.std),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
  addLegend(pal = pal, values = ~en.county$pm25.std, opacity = 0.5, title = 'Daily variability of PM 2.5',
            position = "bottomright")
pm.25_map
```
Discussion of the map:

From the map, we can find out that the areas with lowest PM 2.5 variability (<4.2) in red are located in the southern-east Buffalo areas. And the highest daily variability(>5) of PM 2.5 in green are distributed in the northern-west part of Buffalo, near the Niagara falls and the boundary bewteen the Canada and US. From the northern-west corner to the southern-east corner, the color in the maps turns from green to red, which means the daily variability of PM 2.5 decrease gradually in spatial.

# Question2
Solve the p-median problem when p value is ranged from 3 to 15 stations (p = 3,4,...,15). 
```{r}
set.seed(12345)
distance_total = c(rep(0,13))
set.seed(12345)
for (i in 3:15){
  solution.i <- allocations(en.county, p=i)
  unique(solution.i$allocation)%>%print()
  distance_total[i-2]= sum(unique(solution.i$allocdist))
}
distance_total
p_facility = c(3:15)
TotalDistance = distance_total/100000
par(mar = rep(2, 4))
```
## Report the solutions for each value of p in terms of weighted distance and the sites selected for facilities:

When the p is 3, the total weighted distance is 2971383, and the selected sites are 55, 287, 178.
When the p is 4, the total weighted distance is 2564183, and the selected sites are 18, 17, 284, 178.
When the p is 5, the total weighted distance is 2246062, and the selected sites are 18, 49, 288, 245, 159.
When the p is 6, the total weighted distance is 2045417, and the selected sites are 14, 17, 296, 246, 159, 137.
When the p is 7, the total weighted distance is 1849296, and the selected sites are 14, 49, 225, 275, 88, 159, 124.
When the p is 8, the total weighted distance is 1734819, and the selected sites are 14, 49, 224, 297, 167, 195, 183, 197.
When the p is 9, the total weighted distance is 1658696, and the selected sites are 14, 49, 207, 275, 83, 181, 148, 89, 195.
When the p is 10, the total weighted distance is 1568602, and the selected sites are 14, 49, 13, 276, 210, 156, 195, 125, 104, 197.
When the p is 11, the total weighted distance is 1480025, and the selected sites are 14, 49, 13, 76, 210, 173, 195, 184, 281, 104, 140.
When the p is 12, the total weighted distance is 1415789, and the selected sites are 14, 49, 31, 188, 292, 274, 156, 149, 125, 119, 195, 147.
When the p is 13, the total weighted distance is 1338536, and the selected sites are 14, 49, 13, 188, 76, 210, 156, 181, 184, 281, 119, 195, 140.
When the p is 14, the total weighted distance is 1274066, and the selected sites are 14, 49, 13, 188, 76, 210, 156, 181, 125, 281, 119, 195, 147, 241.
When the p is 15, the total weighted distance is 1232094, and the selected sites are 14, 2, 49, 13, 188, 76, 210, 156, 181, 125, 281, 119, 195, 147, 241

The site: 14, 49, 13, 188, 210, 188, 195, 119, 147, 241 are consistently identified as facilities.

## Plot the tradeoff curve of values,  Y axis for weighted distance and X axis for p
```{r}
ggplot(,aes(p_facility, distance_total))+geom_line() +geom_point()+xlim(0,15)
```

## The map of the spider diagram when p = 6
```{r}
set.seed(12345)
solution.6 <- allocations(en.county1, p=6)
unique(solution.6$allocation)%>%print()
#distance_total6= sum(unique(solution.6$allocdist))
#distance_total6

plot_p6 = plot(star.diagram(solution.6),col='darkblue',lwd=2,add=FALSE)
distance_total6_seperate= unique(solution.6$allocdist)
#distance_total6_seperate

n <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(0,0.02,0.04,0.06,0.08,0.10,0.12,0.14,0.16,0.18,0.2,Inf)
pal<-colorBin("Spectral", domain = solution.6$allocdist, bins = bins)

A=star.diagram(solution.6)


map1 <- n%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(solution.6$allocdist),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
  addLegend(pal = pal, values = ~solution.6$allocdist, opacity = 0.4, title = 'The allocation distance',
            position = "bottomright")%>%
  addPolylines(data=A,
              color = '#F00',
              weight = 2)
map1
```

For the p = 6 solution, the best solution uses sites 14  49 292 167 146 183.

# Question3
For the p = 5 solution, the best solution uses sites 18, 49, 121, 159, 230
Graduate students:

1)Solve the p-median problem using allocations in order to ﬁnd the best solution which does not involve using all ﬁve of these sites 
```{r}
set.seed(12345)
encounty2 = en.county[-c(18,49,121,159,230),]
solution.5_new <- allocations(en.county, encounty2, p =5)
unique(solution.5_new$allocation)
```

2) How does the best 5 facility solution compare to the one that doesnt use all of the best sites, in terms of weighted distance, and in terms of sites used? 


## Distance and sites when use all of the best sites
```{r}
set.seed(12345)
solution.5 <- allocations(en.county,p =5)
unique(solution.5$allocation)
distance_total5= sum(unique(solution.5$allocdist))
distance_total5
```
When we use all the best sites, the sites are 18  49 288 245 159. And the weighted distance are total 2246062.

## Distance and sites when use the sites which does not involve using all ﬁve of these sites: 18, 49, 121, 159, 230
```{r}
set.seed(12345)
encounty2 = en.county[-c(18,49,121,159,230),]
solution.5_new <- allocations(en.county, encounty2, p =5)
unique(solution.5_new$allocation)
distance_total5_new= sum(unique(solution.5_new$allocdist))
distance_total5_new
```
When we do not use all the best sites, the sites are 14   6 225 119 146. And the weighted distance are total 2266776.
It's obvious that the weighted distance not under the case of best sites are much larger than the weighted distance under the best sites.

2) Map these two solutions using the spider diagram and report the id of ﬁve sites for each solution.

## Map when use all of the best sites
```{r}
set.seed(12345)
solution.5_1 <- allocations(en.county1,p =5)
unique(solution.5_1$allocation)
distance_total5_1= sum(unique(solution.5_1$allocdist))
distance_total5_1

n <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(0,0.02,0.04,0.06,0.08,0.10,0.12,0.14,0.16,0.18,0.2,Inf)
pal<-colorBin("Spectral", domain = solution.5_1$allocdist, bins = bins)

B=star.diagram(solution.5_1)

map2 <- n%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(solution.5_1$allocdist),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
  addLegend(pal = pal, values = ~solution.5_1$allocdist, opacity = 0.7, title = 'The allocation distance',
            position = "bottomright")%>%
  addPolylines(data=B,
               color = '#F00',
               weight = 2)
map2
```

## Map when use the sites which does not involve using all ﬁve of these sites: 18, 49, 121, 159, 230
```{r}
set.seed(12345)
encounty2 = en.county1[-c(18,49,121,159,230),]
solution.5_new_1 <- allocations(en.county1, encounty2, p =5)
unique(solution.5_new_1$allocation)
distance_total5_new_1= sum(unique(solution.5_new_1$allocdist))
distance_total5_new_1

n <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(0,0.02,0.04,0.06,0.08,0.10,0.12,0.14,0.16,0.18,0.2,Inf)
pal<-colorBin("Spectral", domain = solution.5_new_1$allocdist, bins = bins)

C=star.diagram(solution.5_new_1)

map3 <- n%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(solution.5_new_1$allocdist),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
  addLegend(pal = pal, values = ~solution.5_new_1$allocdist, opacity = 0.7, title = 'The allocation distance',
            position = "bottomright")%>%
  addPolylines(data=C,
               color = '#F00',
               weight = 2)
map3
```

# Question 5

For the p = 5 solution, ﬁnd the best solution which takes into account the population density of each demand zone. 
Make the spider diagram of the solution with population density as a weight and compare these maps with the solution based on Euclidian distance (above)
```{r}
set.seed(12345)
weight = en.county$pop.den
eucdists<-euc.dists(en.county, en.county)
weighted_distance = matrix(0,297,297)
for (i in 1:297){
  weighted_distance[i,] = weight[i]*eucdists[i,]
}
#weighted_distance
solution.5_weighteddistance = allocations(en.county, p=5, metric = weighted_distance)
unique(solution.5_weighteddistance$allocation)
distance_total5_weighteddistance= sum(unique(solution.5_weighteddistance$allocdist))
distance_total5_weighteddistance
```
The best sites when the populaion density is identified as weight : 9 232 218  74 156.

## Make the spider diagram when consider the effects of weights
```{r}
set.seed(12345)
weight_new = en.county1$pop.den
eucdists_new<-euc.dists(en.county1, en.county1)
weighted_distance_new = matrix(0,297,297)
for (i in 1:297){
  weighted_distance_new[i,] = weight_new[i]*eucdists_new[i,]
}

solution.5_weighteddistance_new = allocations(en.county1, p=5, metric = weighted_distance_new)
unique(solution.5_weighteddistance_new$allocation)

p <- leaflet(en.county1) %>%
  setView(-78.80,42.90, 9) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
bins<-c(0,0.02,0.04,0.06,0.08,0.10,0.12,0.14,0.16,0.18,0.2,Inf)
pal<-colorBin("Spectral", domain = solution.5_weighteddistance_new$allocdist, bins = bins, reverse = TRUE)

k=star.diagram(solution.5_weighteddistance_new)

map4 <- p%>%
  addTiles()%>%
  addPolygons(
    fillColor = ~pal(solution.5_weighteddistance_new$allocdist),
    weight = 0.5,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7)%>%
  addLegend(pal = pal, values = ~solution.5_weighteddistance_new$allocdist, opacity = 0.7, title = 'The allocation distance',
            position = "bottomright")%>%
  addPolylines(data=k,
               color = '#F00',
               weight = 2)
map4
```
For the p = 5 solution, the best solution under the weighted distance uses sites 9, 232, 218, 74, 156.
And the total weighted distance is 2089158, which is lower than the distance without considering the weight. When we compare this map which is made with the weighted distance and the above map which is only depent on the euclidian distance, we can find out the best sites are more concentrated in this map.


